BioTime

### 2019-04-16 biotime.Rmd
Merge biotime and trait data, check coverage.

OTT is the best authority but doesn't have common names, so we use the next best ITIS
```{r}
#Getting matches by scientific name 
biotime <- data %>% 
  #filter(taxa == "Birds") %>%
  select(genus_species) %>%
  drop_na() %>%
  distinct() %>%
  mutate(id = get_ids(genus_species, "itis")) %>%
  left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  drop_na(acceptedNameUsageID)

#There is not a unique taxonID for each species, different pop's have different ID's for some marine mammals
#Names with more than one match that need to be resolved manually
#biotime_ids %>% group_by(sort) %>% filter(n()>1) %>% View()

biotime_data <- biotime_ids %>%
  group_by(sort) %>%
  filter(n() < 2) %>%
  ungroup(sort) %>%
  select(taxonID,input) %>%
  left_join(data, by = c("input" = "genus_species"))
```

>>>>

OTT is the best authority but doesn't have common names, so we use the next best ITIS. We'll start with BioTime data:

##Biotime

```{r}
#Getting matches by scientific name 
biotime <- data %>% 
  #filter(taxa == "Birds") %>%
  select(genus_species) %>%
  drop_na() %>%
  distinct() %>%
  mutate(id = get_ids(genus_species, "itis")) %>%
  left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}
unres <- biotime_ids %>%
  distinct(acceptedNameUsageID, input, sort) %>%
  group_by(sort) %>% 
  filter(n()>1) %>% 
  ungroup()

unres
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join observations back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID's
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

biotime_data <- biotime_ids %>%
  select(id = acceptedNameUsageID,input) %>%
  distinct() %>%
  right_join(data, by = c("input" = "genus_species")) %>%
  rename(genus_species = input)
```

##Elton Traits

```{r}
bird_ids <- by_name(unique(bird_trait$scientific), "itis") %>%
  bind_rows(by_common(unique(bird_trait$scientific), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

check if all bird biotime data w/id's has an elton match. There are 54 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_ids$acceptedNameUsageID) %>%
  dim()
```

and 605 entries in the species column without an ID
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id, genus_species) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```
